### 2016-06-07 paper/icfp16-cameraready/evaluation.tex
misc related work

\subsection{Witness Coverage}
\label{sec:eval:witness-coverage}
%
We ran our search algorithm on each program for 1,000 iterations, with
the entry point set to the function that \ocaml\ had identified as
containing a type error.
%
Due to the possibility of non-termination we set a timeout of 10 minutes
per program.
%
% Due to the possibility of non-termination we set a limit on the number
% of reductions to perform, increasing in 500-step increments from 500
% steps to 3,000 steps total.
%
We also added a na{\"\i}ve check for infinite recursion; at each recursive
function call we check whether the new arguments are identical to the
current arguments.
%
If so, the function cannot possibly terminate and we report an error.
%
While not a \emph{type error}, infinite recursion is still a clear bug
in the program, and thus valuable feedback for the user.

@@

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3--4\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 2--3\% triggered an infinite
recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after 10 minutes.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

>>>>

\subsection{Witness Coverage}
\label{sec:eval:witness-coverage}
%
We ran our search algorithm on each program for 1,000 iterations, with
the entry point set to the function that \ocaml\ had identified as
containing a type error.
%
Due to the possibility of non-termination we set a timeout of one minute
total per program.
%
% Due to the possibility of non-termination we set a limit on the number
% of reductions to perform, increasing in 500-step increments from 500
% steps to 3,000 steps total.
%
We also added a na{\"\i}ve check for infinite recursion; at each recursive
function call we check whether the new arguments are identical to the
current arguments.
%
If so, the function cannot possibly terminate and we report an error.
%
While not a \emph{type error}, infinite recursion is still a clear bug
in the program, and thus valuable feedback for the user.

@@

\paragraph{Results}
\label{sec:results-witness}
The results of our experiments are summarized in
Figure~\ref{fig:results-witness}.
%
In both datasets our tool was able to find a witness for 83\% of the
programs in under one second, \ie\ fast enough to be integrated as a
compile-time check. If we extend our tolerance to a 10 second timeout,
we hit a maximum of 87\% coverage.
%
Interestingly, while the vast majority of witnesses corresponded to a
type-error, as expected, 3--4\% triggered an unbound variable error (even
though \ocaml\ reported a type error) and 2--3\% triggered an infinite
recursion error.
%
For the remaining 12\% of programs we were unable to provide any useful
feedback as they either completed 1,000 tests successfully, or timed out
after one minute.
%
% XX programs were deemed safe and XX timed out even at 3,000 steps, \ie
% we could not provide any useful feedback for XX\% of the total programs.
%
While a more advanced search procedure, \eg\ dynamic-symbolic execution,
could likely trigger more of the type errors, our experiments show that
type errors are coarse enough (or that novice programs are \emph{simple}
enough) that these techniques are not necessary.

### 2016-06-07 paper/icfp16-cameraready/related.tex
misc related work

Lerner \etal~\cite{Lerner2007-dt} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{Chen2014-gd} use a variational type
system to allow for the possibility of changing an expression's type,
and search for an expression whose type can be changed such that type
inference would succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types in a program.
%
Neubauer and Thiemann~\cite{Neubauer2003-xv} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{Zhang2014-lv} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{Pavlinovic2014-mr} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis \etal~\cite{Vytiniotis2012-gh} extend the Haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.

@@

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} -- containing holes -- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera's system focuses on backward exploration; in
contrast, our visualization supports forward \emph{and} backward
exploration, though our backward steps are more limited than
Perera's. Specifically, we do not support selecting a value and
inserting the intermediate terms that preceded it while ignoring
unrelated computation steps; this would be interesting future work.

>>>>

Lerner \etal~\cite{Lerner2007-dt} attempt to localize the type
error and suggest a fix by replacing expressions (or removing them
entirely) with alternatives based on the surrounding program context.
%
Chen and Erwig~\cite{Chen2014-gd} use a variational type
system to allow for the possibility of changing an expression's type,
and search for an expression whose type can be changed such that type
inference would succeed.
%
% It then attempts to deduce the error source by searching for an
% expression whose type can be changed such that type inference would
% succeed.
%
In contrast to Lerner, who searches for changes at the
value-level, by searching at the type level Chen's search is complete due
the finite universe of types used in the program.
%
Neubauer and Thiemann~\cite{Neubauer2003-xv} present a
decidable type system based on discriminative sum types, in which all
terms are typeable and type derivations contain all type errors in a
program. They then use the typing derivation to slice out the parts of
the expression related to each error.
%
Zhang and Myers~\cite{Zhang2014-lv,Zhang2015-yu} present an algorithm for
identifying the most likely culprit in a system of unsatisfiable
constraints (\eg type equalities), based on Bayesian reasoning.
%
Pavlinovic \etal~\cite{Pavlinovic2014-mr} translate the %error
localization problem to a MaxSMT optimization problem, using
compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}

@@

\paragraph{Running Ill-Typed Programs}
\label{sec:running-ill-typed}
Vytiniotis~\etal~\cite{Vytiniotis2012-gh} extend the \haskell
compiler GHC to support compiling ill-typed programs, but their intent
is rather different from ours. Their goal was to allow programmers to
incrementally test refactorings, which often cause type errors in
distant functions. They replace any expression that fails to type
check with a \emph{runtime} error, but do not check types
at runtime.
%
Bayne~\etal~\cite{Bayne2011-cn} also provide a semantics for running
ill-typed (\java) programs, but in constrast transform the program to
perform nearly all type checking at run-time. The key difference between
Bayne~\etal and our work is that we use the dynamic semantics to
automatically search for a witness to the type error, while their focus
is on incremental, programmer-driven testing.

@@

\paragraph{Program Exploration}
Perera \etal~\cite{Perera2012-dy} present a tracing semantics
for functional programs that tags values with their provenance, enabling
a form of backwards program slicing from a final value to the sequence
of reductions that produced it. Notably, they allow the user to supply a
\emph{partial value} --- containing holes --- and present a partial slice,
containing only those steps that affected the the partial value.
% This
% system is designed to answer questions of the form ``Where did this
% value come from?'' and thus is focused on backward exploration. 
Perera's system focuses on backward exploration; in
contrast, our visualization supports forward \emph{and} backward
exploration, though our backward steps are more limited than
Perera's. Specifically, we do not support selecting a value and
inserting the intermediate terms that preceded it while ignoring
unrelated computation steps. %; this would be interesting future work.
